### 2024-05-14 02_codes/paper_draft.qmd
text

---
title: "Do Non-Cognitive Skills Affect Employment Outcomes During the School-to-Work Transition?"
author: "Garen Avanesian, Marina Borovskaya, Valery Egorova, Marina Masych, Lyudmila Dikaya"
format: pdf
editor: visual
bibliography: references.bib
---

<https://www.ilo.org/publications/characterizing-school-work-transitions-young-men-and-women-evidence-ilo>

https://academic.oup.com/workar/article/9/3/239/7140451

<https://www.ilo.org/resource/school-work-transition-survey-swts>

## Introduction

The transition from school to work represents a pivotal phase in the lives of young individuals, marked by decisions that significantly shape their future careers and overall lives. This transition is influenced by various factors such as education, training, as well as supply and demand for qualified workforce on the labor market. Recent studies point out that non-cognitive skills refer to one of the critical components in this transition [@lerman2013; @glewwe2017; @zudina2022; @ripamonti2023; @avanesian2024]. These skills encompass characteristics of personality not directly tied to cognitive abilities, including motivation, self-control, and social skills. Research increasingly suggests that non-cognitive skills play a crucial role in determining success in the labor market, influencing both employment opportunities and earning potential.

## Data

descriptive statistics table

age, sex, area of residence, ses quintile, edu level

## Econometric Verification Strategy

## Results

### Non-Cognitive Skills of Youth

Here we need a table or a chart with differences in NCS by employment status

### Socio-Demographic Profile of Employed Youth

### Effect of Non-Cognitive Skills on Employment Outcomes

We need to have two mlm models

1\)

## Discussion

### Research Limitations

## Policy Implications

## Conclusion

### 2024-05-15 02_codes/paper_draft.qmd
text

@@
 ---
-title: "Do Non-Cognitive Skills Affect Employment Outcomes During the School-to-Work Transition?"
+title: "Do Non-Cognitive Skills Affect Employment Outcomes of Youth During the School-to-Work Transition?"
 author: "Garen Avanesian, Marina Borovskaya, Valery Egorova, Marina Masych, Lyudmila Dikaya"
@@
 
+Research questions
+
+1\) do non-cognitive skills affect the probability of employment for youth during school-to-work transition? If yes, what are the most influential skills?
+
+2\) does this effect is heterogeneous with respect to socio-economic status? in other words, which non-cognitive skills are the most important when it comes to the youth from the poorest households
+
 ## Data
@@
 
+The major issue with respect to predict effects of non-cognitive skills on employment outcomes of youth during school-to-work transition refers to unobserved heterogeneity present due to the multiple factors.
+
+1\) The age range of the study is 15-29 in accordance with the ILO definition of the school-to-work transition.
+however we understand that there are many factors that might not be captured by the model but which affect baseline probability of employment due to age.
+while we can account for level of completed education, there could be a number of issues in terms of labor market entry for young workforce, including young professionals.
+
+2\) Russia is a big federal state, and there is a number of unobserved, systemic factors of labor demand and supply that affect employment at the regional level.
+In other words, country regions establish huge heterogeneity in employment of youth.
+
+3\) Finally, existing structures of socio-economic inequality make substantial footprint on school-to-work transition of young people, with youth from the richer households having access to better schools, parents being included in professional networks , which results in having more default options for youth to find a job.
+with that regard, the poor , who experience greater risks of lacking access to social lifts to elevate their status in the society, might also experience higher risks of unemployment either due to the lower human capital they possess or due to other factors.
+with that regard, for the purpose of labor policy it is essentially important to understand which non-cognitive skills could help economically disadvantaged youth have access to jobs.
+
+These issues can be well resolved by the multilevel (mixed-effect) regression model.
+
+describe briefly what it is and what are the advantages of this model in the context of research question.
+
+we built two mixed effect models to address the research questions.
+
+the first one regresses binary variable employment on (fixed part) sex, area of residence, parental education, parental occupation (?), NCS
+
 ## Results
@@
 
-### Socio-Demographic Profile of Employed Youth
+### Socio-Demographic Profile of Employed Youth - ???
 

### 2024-06-18 02_codes/paper_draft.qmd
text and folder structure

@@
 
+*Masych, Dikaya, Avanesian*
+
 The transition from school to work represents a pivotal phase in the lives of young individuals, marked by decisions that significantly shape their future careers and overall lives.
@@
 
-Research questions
+Нужно дать обзор литературы, в которой некогнитивные навыки и личностные характеристики влияют на трудоустройство, особенно с прицелом на молодежь
 
-1\) do non-cognitive skills affect the probability of employment for youth during school-to-work transition? If yes, what are the most influential skills?
+Самая ранняя статья Рожковой (Вопросы Экономики) про отдачу от неекогнитивных навыков содержит блок про трудоустройство
 
-2\) does this effect is heterogeneous with respect to socio-economic status? in other words, which non-cognitive skills are the most important when it comes to the youth from the poorest households
+То же самое касается препринта Гимпельсона, Зудиной и Капелюшникова.
+
+Это все примеры про Россию.
+
+Мы также можем сдкелать ссылки на статью Зудиной по НИТ, и нашу статью.
+
+Но примеры нельзя ограничивать Россией и нужна хорошая международная литература - частично она уже упомянута в нашей статье.
+
+Где-то во введении нужно также дать контекст и посмотреть данные Росстата по молодежной безработице в России, где из регионов она самая низкая и где самая высокая.
+Если молодежной безработицы нет, то можно просто безработицу.
+Несколько потенциальных источников
+
+<https://cyberleninka.ru/article/n/molodezhnaya-bezrabotitsa-v-rossii-prichiny-i-puti-snizheniya/viewer>
+
+<https://vital.lib.tsu.ru/vital/access/services/Download/vtls:000587153/SOURCE1>
+
+Не нужно много о безработице, но необходимо дать контекст
+
+Current study explores the impact of non-cognitive skills on employment outcomes of youth aged 15-29 during school-to-work transition in Russia.
+Based on data of Russia Longitudinal Monitoring Survey (RLMS), it aims to address the following research question: do non-cognitive skills affect the probability of employment for youth during school-to-work transition?
+If yes, what are the most influential skills?
+The consideration of this research question is guided by supplementary questions.
+First of all, the study aims to explore if the effect of non-cognitive skills on employment outcones is heterogeneous with respect to socio-economic status of an individual.
+in other words, which non-cognitive skills are the most important when it comes to the youth from the poorest households
+
+-   calculate random slope terms for different completed education levels (make a separate regression model for each education level)
 
 ## Data
+
+*Egorova*
+
+RLMS data for 2019 is to be used. Sample 15-29 years old as it aligns with ILO definition for school-to-work transition.
 
@@
 
-## Econometric Verification Strategy
+## Methodology
+
+*Avanesian*
 
@@
 
-### Non-Cognitive Skills of Youth
+### Non-Cognitive Skills of Youth (NCS)
+
+*Egorova*
 
@@
 
-### Socio-Demographic Profile of Employed Youth - ???
-
 ### Effect of Non-Cognitive Skills on Employment Outcomes
 
-We need to have two mlm models
+*Avanesian*
 
-1\)
+We need to have three regression models
 
@@
 
+*Masych, Dikaya*
+
 ### Research Limitations
+
+*Avanesian*
 
@@
 
+*Masych*
+
 ## Conclusion
+
+*Avanesian*
+
+## References

### 2024-06-24 02_codes/paper_draft.qmd
[data merge]

@@
 The consideration of this research question is guided by supplementary questions.
-First of all, the study aims to explore if the effect of non-cognitive skills on employment outcones is heterogeneous with respect to socio-economic status of an individual.
+First of all, the study aims to explore if the effect of non-cognitive skills on employment outcomes is heterogeneous with respect to socio-economic status of an individual.
 in other words, which non-cognitive skills are the most important when it comes to the youth from the poorest households

### 2024-06-24 02_codes/0200_load_packages.R

@@ -1,0 +1,46 @@
+#-------------------------------------------------------------------
+# Project: Scientific Involvement Indincator
+# Organization: SFedU Future Skills Research Lab
+# Objective: Install packages needed for the analysis
+# Author: Valeria Egorova
+# Date: 19 Nov 2023
+#-------------------------------------------------------------------
+
+
+# List of required packages
+
+required_packages <- c("tidyverse",
+                       "haven",
+                       "readxl",
+                       "psych",
+                       "foreign",
+                       "factoextra",
+                       "corrr",
+                       "ggcorrplot",
+                       "writexl",
+                       "car",
+                       "ggfortify",
+                       "mgcv",
+                       "erer",
+                       "caret",
+                       "e1071",
+                       "margins",
+                       "Hmisc",
+                       "naniar",
+                       "ggpubr",
+                       "DescTools")
+
+
+# Function to check and install packages
+check_and_install_packages <- function(packages) {
+  for (package in packages) {
+    if (!requireNamespace(package, quietly = TRUE)) {
+      install.packages(package)
+    }
+    library(package, character.only = TRUE)
+  }
+}
+
+
+# Call the function with the list of required packages
+check_and_install_packages(required_packages)

### 2024-06-24 02_codes/0201_data_merge.R

@@ -1,0 +1,136 @@
+#-------------------------------------------------------------------
+# Project: Do Non-Cognitive Skills Affect Employment Outcomes of Youth During the School-to-Work Transition?
+# Organization: SFedU Future Skills Research Lab
+# Objective: Prepare a dataset for the analysis
+# Author:  Garen Avanesian, Valery Egorova
+# Date: 24 June 2024
+#-------------------------------------------------------------------
+
+source(file.path(rcodes, "0200_load_packages.R"))
+
+non_cogn  <- read_sav("01_input_data/r28i_open_2023.sav")
+hi        <- read_sav("01_input_data/r28iall_61.sav")
+hh        <- read_sav("01_input_data/r28hall_61.sav")
+
+data_hh <-
+  hh %>%
+  select(xb1.o, xf14, xa3, xredid_h, xid_h) %>%
+  replace_with_na(replace = list(xf14 = c(99999997, 99999998, 99999999))) %>%
+  mutate(x = xf14/xb1.o,
+         number_of_fmem = xb1.o,
+         windex = percent_rank(x),
+         Income = xf14,
+         family = xa3,
+         windex5 = case_when(windex <= 0.20 ~ 1,
+                             windex > 0.20 & windex <= 0.40 ~ 2,
+                             windex > 0.4 & windex <= 0.60 ~ 3,
+                             windex >= 0.61 & windex <= 0.80 ~ 4,
+                             windex > 0.8  ~ 5,
+                             T ~ as.numeric(NA))) %>%
+  select(windex5, family, Income, windex, number_of_fmem, xredid_h, xa3, xid_h, windex, x)
+
+data_hi <- 
+  hi %>%
+  select(idind, xredid_i, status, xh5, xj1, x_age, xid_h,  xj161.3y, xj161.3m,
+         xj70.2, xj72.1a, xj72.2a, xj72.3a, xj72.4a, xj72.5a, xj72.6a, xj81, 
+         x_inwgt, x_diplom, xh3, xj65, xj62, xj322, xj72.171, xm3, xj13.2) %>%
+  replace_with_na_all(condition ~.x %in% c(99999996, 99999997, 99999998, 99999999)) %>%
+  mutate(area = case_when(status %in% c(1,2) ~ "urban",
+                         status %in% c(3,4) ~ "rural",
[…]

### 2024-06-24 user_profile.R

@@ -1,0 +1,39 @@
+#-------------------------------------------------------------------
+# Project: Do Non-Cognitive Skills Affect Employment Outcomes of Youth During the School-to-Work Transition
+# Organization: SFedU Future Skills Research Lab
+# Author:  Garen Avanesian, Valery Egorova
+# Date: 24 June 2024
+#-------------------------------------------------------------------
+
+
+USERNAME    <- Sys.getenv("USERNAME") 
+USER        <- Sys.getenv("USER")
+
+
+if (USERNAME == "Valery") {   
+  projectFolder  <- getwd()   
+}
+
+if (USER == "karavan88") {   
+  projectFolder  <- getwd()   
+}
+
+
+stopifnot(dir.exists(projectFolder))
+
+# set up key folders
+documentation <-  file.path(projectFolder, "00_documentation")
+inputData     <-  file.path(projectFolder, "01_input_data")
+rcodes        <-  file.path(projectFolder, "02_codes")
+outData       <-  file.path(projectFolder, "03_outputs/0301_data")
+outTables     <-  file.path(projectFolder, "03_outputs/0302_tables")  
+outFigures    <-  file.path(projectFolder, "03_outputs/0303_figures")
+
+
+stopifnot(dir.exists(documentation))
+stopifnot(dir.exists(inputData))
+stopifnot(dir.exists(rcodes))
+stopifnot(dir.exists(outData))
+stopifnot(dir.exists(outTables))
+stopifnot(dir.exists(outFigures))
+

### 2024-09-26 02_codes/paper_draft.qmd
[text]

--- 2024-05-15 02_codes/paper_draft.qmd
@@
 title: "Do Non-Cognitive Skills Affect Employment Outcomes of Youth During the School-to-Work Transition?" 
-author: "Garen Avanesian, Marina Borovskaya, Valery Egorova, Marina Masych, Lyudmila Dikaya"
+author: "Garen Avanesian, Southern Federal University"
 format: pdf
@@
 bibliography: references.bib
+biblio-style: bwl-FU
+citeproc: true
+csl: "apa-with-abstract.csl"
 ---
 
-<https://www.ilo.org/publications/characterizing-school-work-transitions-young-men-and-women-evidence-ilo>
+**Abstract**
 
-https://academic.oup.com/workar/article/9/3/239/7140451
+The study analyses the impact of non-cognitive skills on employment outcomes of youth aged 15-29 during the school-to-work transition in Russia.
+Based on the data of the Russian Longitudinal Monitoring Survey collected in 20XX.
+Adopting a mixed-effects model approach, the study aims to estimate if the effect of non-cognitive skills on employment is heterogeneous due to the income level and completed level of education.
+The study finds out that......
 
-<https://www.ilo.org/resource/school-work-transition-survey-swts>
+**Key words:** non-cognitive skills, school-to-work transition, human capital, personality traits, big five, socio-economic status, mixed-effects models
+
+**JEL Codes:** I24, J24; C51
+
+\newpage
 
@@
 The transition from school to work represents a pivotal phase in the lives of young individuals, marked by decisions that significantly shape their future careers and overall lives.
-This transition is influenced by various factors such as education, training, as well as supply and demand for qualified workforce on the labor market.
-Recent studies point out that non-cognitive skills refer to one of the critical components in this transition [@lerman2013; @glewwe2017; @zudina2022; @ripamonti2023; @avanesian2024].
+Importance of school-to-work transition is emphasized by the fact that it has long-term implications on socio-economic outcomes of individuals beyond youth period: while success in getting a first job that matches skills and expectations constitutes the further path to occupational success, early non-employment or poor transition to the world of work have long-lasting and negative implications on professional outcomes and beyond [@akkermans2021; @zacher2021; @baert2013; @luijkx2009; @verbruggen2015].
+Moreover, the evidence points out that "association between having a high level of education and securing permanent full-time employment is weaker than expected" and level of education does not lead to the higher levels of job satisfaction [@chesters2020].
+While there are many approaches to the definitions of school-to-work transition, their extensive survey is provided in [@blokker2023].
+In this work, I rely on the one that is provided by ILO and defines it "as the passage of a young person (aged 15 to 29 years13) from the end of schooling to the first fixed-term or satisfactory employment" [@ilo2010, p.
+4].
+With that respect, each young person aged 15-29 is classified under the following categories: (1) transition not yet started for those who are either in school or not in school, inactive and with no intention to look for a job; (2) in transition, for those who are either unemployed, or in unsatisfactory self-employment temporary job, in unpaid family employment, or not in school but searching for a job; (3) transited, for those who are either in fixed-term employment or in satisfactory temporary or self-employment.
+While Sustainable Development Goals define youth as the population in the age 15 to 24 yeras old, ILO extended the upper bound to acknowledge that many young women and men remain in education till 24 years, and a wider age frame is required to capture employment experiences of youth post graduation.
+
+The success of school-to-work transition, i.e., if a young person finds a decent job, is determined by various factors such as education, training, as well as supply and demand for qualified workforce on the labor market.
+Those young people who, in addition to being unemployed, also do not study, refer to the most vulnerable group in the school-to-work transition and are defined as not in education, employment, or training (NEET).
+In the recent past, two events amplified vulnerability of youth in terms of labor market outcomes: the Great recession of 2009 and COVID-19 outbreak.
+The evidence points out that the financial crisis substantially challenged the integration of young people on the labor market all over the world [@tanveerchoudhry2012; @kelly2015; @verick2011; @verd2019; @montalvao2017].
+Young people generally and youth employment particularly were put at the center of the global sustainable development agenda, summarized in Sustainable Development Goal 8 on decent jobs and economic growth.[^1] The global pandemic happened in 2019 amplified the labor market challenges faced by the global community.
+It "has heightened the level of uncertainty experienced by many young people worldwide and has complicated the developmental tasks associated with the transition to adulthood, including leaving home, completing education, and obtaining full-time employment" [@allmang2022].
+The initial pandemic estimates suggested that young people were hit more than others by job loss, fell in working hours and substantial drops of income [@ilo2020youth].
+It was estimated that as the result of the pandemic shock in 2020 globally youth employment rates fell by 8.7% in comparison to a 3.7 % drop for adults [@ilo2021].
+
+[^1]: Namely, SDG8.6, which asks countries to report the share of youth Not in Education, Employment, or Training.
+
+Though initially the problems of employment and labor market were in the primary interest of economists, the analysis of available academic literature illustrates that in the recent body of work, school-to-work transition was primarily in the focus of psyhologists, sociologists, and then economists, which marks the interdisciplinary nature of the topic [@blokker2023].
+With different disciplines bringing various aspects of employment that were not conventionally regarded by economists, the recent studies point out that non-cognitive skills refer to one of the critical components in school-to-work transition [@lerman2013; @glewwe2017; @zudina2022; @ripamonti2023; @avanesian2024].
 These skills encompass characteristics of personality not directly tied to cognitive abilities, including motivation, self-control, and social skills.
-Research increasingly suggests that non-cognitive skills play a crucial role in determining success in the labor market, influencing both employment opportunities and earning potential.
+Research increasingly suggests that non-cognitive skills play a crucial role in determining success in the labor market, influencing both employment opportunities and earning potential [@ferguson2011; @almlund2011a; @almlund2011].
 
-Research questions
+While the debates on the taxonomy of non-cognitive skills are still present in the literature, with different scholars using different characteristics in measuring their impact on socio-economic outcomes, a five-factor model of personality traits, also known as "Big Five Inventory" [@goldberg1990], became the most mainstream approach in economics.
+Under this framework, all the variations in personality can be summarized by five orthogonal and independent characteristics, which include openness to new experience, conscientiousness, extraversion, agreeableness, and neuroticism (emotional stability).
+The state of the art research in economics, psychology, and sociology points out that these personality factors are pivotal in determining the outcomes of youth during school-to-work transition.
+Evidence suggests that personality measured by big five taxonomy makes a strong impact on wellbeing of youth during the transition from graduate schools to the world of work [@buhl2007].
+Another study based on analysis of the longitudinal data of 685 Dutch vocational training graduates showed that "extraversion and emotional stability were related to better job-search outcomes after graduation", and "some relations between Big Five personality traits and job-search outcomes were explained by social capital" regarded as "available resources through social relations" [@baay2014, p.739].
 
-1\) do non-cognitive skills affect the probability of employment for youth during school-to-work transition? If yes, what are the most influential skills?
+A study that used the data on participants from early childhood to retirement, assessed the impact of personality traits across the life span on intrinsic (job satisfaction) and extrinsic career success (occupation status and income) [@judge1999].
+It found out that conscientiousness positively predicts both intrinsic and extrinsic career success, while neuroticism negatively predicts extrinsic success.
+Overall, controlling for general cognitive ability, the study identified that even measured in early childhood, personality traits explain substantial portion of variance in career outcomes.
+Based on the data of the German Socio-Economic Panel (SOEP), another study estimated the effect of five personality factors on the duration of unemployment.
+The results highlight that such traits as conscientiousness and neuroticism have a strong impact on the instantaneous probability of finding a job, with the first affecting it positively, whereas the latter producing a negative effect [@uysal2011].
 
-2\) does this effect is heterogeneous with respect to socio-economic status? in other words, which non-cognitive skills are the most important when it comes to the youth from the poorest households
+A few studies were exploring the role of non-cognitive skills measured by BFI with respect to NEET youth, the most vulnerable cohort in the school-to-work transition.
+Based on Russian Longitudinal Monitoring Survey, they produced consistent findings that point out at the negative association between conscientiousness and NEET status, and positive relationship between neuroticism and NEET status [@avanesian2024; @zudina2022].
+
+Another research pillar explored the effect of non-cognitive skills on school-to-work transition beyond Big Five personality traits.
+The large body of work indicates the special role of self-efficacy in the school-to-work transition [@lent1999; @emirza2021; @tolentino2018; @masdonati2021; @grosemans2018].
+The concept was introduced by psychologist A.
+Bandura who made it central in his social cognitive theory and defined it as the mechanism of personal agency that "refers to beliefs in one’s capabilities to organize and execute the courses of action required to manage prospective situations" [@self-eff1995, p.
+2].
+To put it simpler, this non-cognitive characteristics refers to personal belief in achieving a certain task or accomplishing a goal.
+With that respect, being convinced in the success of getting the job does not guarantee that a person will get it, but it produces the behaviors, actions and attitudes that in the end facilitate achievement of this goal.
+In that light, this concept is somewhat similar to another psychological concept of locus of control, as both are the expressions of personal agency.
+Defined as an extent to which people believe to have control of events in their life as opposed to external forces, this non-cognitive characteristic may partially help to compensate for socio-economic disadvantage regarding avoidance of economic inactivity and unemployment to some extent; however, it did not prove to provide protection against long-term economic inactivity, i.e.
+if an individual spent more than 6 months spent not in education, employment or training [@ng-knight2017].
+
+Current study explores the effect of non-cognitive skills on employment outcomes of youth aged 15-29 during school-to-work transition in Russia.
+The Russian data highlights that youth has the highest unemployment rates: in 2022, they accounted for 26% for 15-19 year olds, 13.6% for 20-24 year olds, and 5.2% for 25-29 year olds.
+This is compared to 2.8% of unemployment for individuals in the age of 40 to 44 year olds, the lowest value in the sample [@rosstat2024, p.
+122].
+Further, out of all higher education graduates between 2020-2022 who find a job, for 24% the obtained employment does not match the received profession.
+Finally, the available national data evidence that the labor market in the country demonstrates unprecedented spatial segregation: with the 4% as a country's average unemployment in 2022, it varies in the range between 1.6% in Yamalo-Nenets Autonomous Okrug and 30% in Republic of Ingushetia.[^2]
+
+[^2]: The Rosstat data were accessed on 24 September 2024, via the following link: <https://rosstat.gov.ru/storage/mediabank/Trud_3_15-s.xlsx>
+
+Based on data of Russia Longitudinal Monitoring Survey (RLMS), this study aims to address the following research question: do non-cognitive skills affect the probability of employment for youth during school-to-work transition?
+If yes, what are the most influential skills?
+The consideration of this research question is guided by supplementary questions.
+First of all, it aims to estimate which age in the school-to-work transition span is associated with the highest probability of getting a job.
+Further, the study explores if the effect of non-cognitive skills on employment outcomes is heterogeneous with respect to socio-economic status (SES) of an individual.
+In other words, which non-cognitive skills are the most important when it comes to the youth from the poorest households?
+This question could have potential implications on the labor market policy in terms of providing the young poor the channels of social mobility.
+Lastly, it is possible to hypothesize that different completed levels of education could produce heterogeneous returns on non-cognitive characteristics when it comes to the employment.
+In other words, the study wants to investigate if the higher levels of education completed are associated with the stronger effects of non-cognitive skills on employment.
 
@@
 
-descriptive statistics table
+The data used in this study is comprised of two waves of Russia Longitudinal Monitoring Survey, the one in 2016 and 2019 as these years correspond to the ones when information on Big Five personality traits was collected.
+Sample those who were 15-29 years old in 2016 was extracted, which resulted in the fact that some of those respondents present in the 2016 wave were 33 years old in 2019.
+This was done for the completeness of the sample.
+It needs to be noted that the selected variables on non-cognitive skills substantially suffered from the missingness.
+If a respondent was present in both waves and responded provided a needed information in at least one of the waves, this value was used to impute and use in the estimation.
+This was done based on the assumption that personality traits are stable in the adulthood (ref).
+The final data accounted for XXXX observations in total, XXXX coming from the 2016 wave and XXXX from the 2019 wave.
+Out of them, XXXX observations were present in both waves.
+The summary statistics are presented on Table 1.
 
-age, sex, area of residence, ses quintile, edu level
+## Methodology
 
-## Econometric Verification Strategy
+In order to estimate the effect of non-cognitive skills on employment outcomes, the study relies on the unbalanced panel, where substantial share of observations has only one record.
+In order to address the research questions that drive this investigation, mixed-effects regression model is adopted.
+It's primary advantage refers to the possibility to control endogeneity that occurs due to the unobserved heterogeneity at multiple levels.
 
-The major issue with respect to predict effects of non-cognitive skills on employment outcomes of youth during school-to-work transition refers to unobserved heterogeneity present due to the multiple factors.
+First, as the data is longitudinal and XX% of the sample has two records in time, the model adopts random intercept ID to separate between- and within-individual variation in employment and non-cognitive skills.
+Further, the age range of the study is 15-29 in accordance with the ILO definition of the school-to-work transition, but due to differences in baseline probability due to age, it would be the source of endogeneity at the individual level.
+Younger people, due to school enrollment, would be associated with low incidence rates of employment.
+With that regard, the model adopted to estimate the effect of non-cognitive skills on employment will have to isolate the effect of age.
+For that, the model adopts the random intercept for age, which allows the probability of employment to vary across age.
 
-1\) The age range of the study is 15-29 in accordance with the ILO definition of the school-to-work transition.
-however we understand that there are many factors that might not be captured by the model but which affect baseline probability of employment due to age.
-while we can account for level of completed education, there could be a number of issues in terms of labor market entry for young workforce, including young professionals.
+Another source of endogeneity would refer to the spatial factors.
+Russia is a big and populated federal state, with explicit spatial disparities in access to the labor market.
+Unobserved and not included into the estimation, these factors inevitably affect a baseline probability of career entry, with the individuals sharing the same individual characteristics having different chances of getting a job purely due to the regional segregation of the labor market.
+In other words, within- and between-regional variation need to be separated in the model, which is accomplished through the random intercept term for region.
 
-2\) Russia is a big federal state, and there is a number of unobserved, systemic factors of labor demand and supply that affect employment at the regional level.
-In other words, country regions establish huge heterogeneity in employment of youth.
+These considerations are reflected in the main model of the study, that in addition to the outlined random terms in predicting employment through Big Five personality traits also controls for sex, area of residence, year of survey, income per capita percentile group as a proxy for SES, current education attendance and highest level of education completed.
+However, two supplementary research questions are reflected in two additional mixed effects models carried out in this study.
 
-3\) Finally, existing structures of socio-economic inequality make substantial footprint on school-to-work transition of young people, with youth from the richer households having access to better schools, parents being included in professional networks , which results in having more default options for youth to find a job.
-with that regard, the poor , who experience greater risks of lacking access to social lifts to elevate their status in the society, might also experience higher risks of unemployment either due to the lower human capital they possess or due to other factors.
-with that regard, for the purpose of labor policy it is essentially important to understand which non-cognitive skills could help economically disadvantaged youth have access to jobs.
+Needs to be taken into account that existing structures of socio-economic inequality make substantial footprint on school-to-work transition of young people, with youth from the richer households having access to better schools, parents being included in networks , which results in having more default options for youth to find a job.
+With that regard, the poor, who experience greater risks of lacking access to social lifts to elevate their status in the society, might also face higher risks of unemployment due to the lower human capital they possess.
+With that regard, for the purpose of labor policy it is essentially important to understand which non-cognitive skills could help economically disadvantaged youth have access to jobs.
+For that, a second model adopts the random slope terms of non-cognitive skills by the income percentile groups.
 
-These issues can be well resolved by the multilevel (mixed-effect) regression model.
-
-describe briefly what it is and what are the advantages of this model in the context of research question.
-
-we built two mixed effect models to address the research questions.
-
-the first one regresses binary variable employment on (fixed part) sex, area of residence, parental education, parental occupation (?), NCS
+Finally, different levels of education would also result in different likelihoods of having a job.
+It also can be assumed that non-cognitive skills might affect differently the probability of employment for those with and without higher education.
+As this effect could be heterogeneous, the verification model should take it into account through the random slope term of the Big Five traits by the highest level of education completed.
+This is done in the third calculated model.
+The data analysis for the study was carried out in R [@R], an open-source software for the statistical computing.
+The mixed-effects regression models were computed using the lme4 package [@bates2015].
 
@@
 
-### Non-Cognitive Skills of Youth
+### Differences in Non-Cognitive Skills
 
-Here we need a table or a chart with differences in NCS by employment status
-
-### Socio-Demographic Profile of Employed Youth - ???
+Box plot with three facets: employment status, Income percentile, Higher Education
 
@@
 
-We need to have two mlm models
+Baseline regression model
 
-1\)
+### Skills Produce Different Effects by Socio-Economic Status
+
+### Does Level of Education Increase the Effect of Non-Cognitive Skills?
 
@@
 
-### Research Limitations
+## Research Limitations
+
+1\) Quality of data due to the misingness and attrition makes the data potentially highly prone to sample selection bias.
+
+Based on the provided methodology section, here are potential research limitations and suggestions to address them.
+I've considered the mentioned data missingness and sample selection bias, as well as other potential issues:
+
+1.  **Missing Data and Sample Selection Bias:**
+
+    -   *Limitation*: The data suffers from missingness, which can lead to sample selection bias.
+
+    -   *Suggestion*:
+
+        -   Implement multiple imputation techniques to handle missing data.
+This involves creating multiple datasets with imputed values for missing observations and then combining the results from analyzing these datasets.
+R packages like 'mice' or 'Amelia' can be used for this purpose.
+
+        -   Perform sensitivity analyses to assess the impact of different missing data handling strategies on the results.
+
+        -   Consider using inverse probability of censoring weighting (IPCW) or other techniques to address sample selection bias due to missing data.
+
+2.  **Unbalanced Panel and Attrition:**
+
+    -   *Limitation*: The panel is unbalanced, with a substantial share of observations having only one record, and there might be attrition bias.
+
+    -   *Suggestion*:
+
+        -   Investigate the reasons for attrition and compare the characteristics of individuals who dropped out with those who remained in the panel.
+
+        -   Consider using techniques like inverse probability of attrition weighting (IPAW) or shared frailty models to account for attrition bias.
+
+3.  **Endogeneity:**
+
+    -   *Limitation*: The study acknowledges endogeneity due to unobserved heterogeneity at multiple levels (age, region, etc.).
+
+    -   *Suggestion*:
+
+        -   While mixed-effects models help control for endogeneity, consider using instrumental variable (IV) approaches or difference-in-differences (DID) methods for specific research questions where suitable instruments or comparison groups can be identified.
+
+        -   Collect more detailed information in future studies to account for additional unobserved factors.
+
+4.  **Measurement Error:**
+
+    -   *Limitation*: Self-reported data on non-cognitive skills and employment outcomes might be subject to measurement error and social desirability bias.
+
+    -   *Suggestion*:
+
+        -   Consider using objective measures of non-cognitive skills (e.g., cognitive tests, observational studies) or validated scales in future research.
+
+        -   Collect data from multiple sources (e.g., respondents and their parents or teachers) to triangulate information.
+
+5.  **Longitudinal Analysis:**
+
+    -   *Limitation*: The study uses a longitudinal dataset but doesn't explicitly address time-varying covariates or dynamic effects.
+
+    -   *Suggestion*:
+
+        -   Incorporate time-varying covariates in the mixed-effects models to capture changes in individual characteristics over time.
+
+        -   Consider using growth curve models or other advanced longitudinal techniques to analyze dynamic effects.
+
+6.  **Interactions and Non-linear Effects:**
+
+    -   *Limitation*: The study focuses on linear effects and interactions, but non-cognitive skills might have non-linear effects on employment outcomes.
+
+    -   *Suggestion*:
+
+        -   Include quadratic or polynomial terms for non-cognitive skills in the models to capture non-linear effects.
+
+        -   Consider using machine learning techniques or non-parametric approaches to model complex relationships.
+
+By addressing these limitations, you can strengthen the validity and robustness of your research findings.
 
@@
 ## Conclusion
+
+## References

### 2024-09-26 02_codes/0201_data_merge3.R
[data merge]

@@ -1,0 +1,32 @@
+#-------------------------------------------------------------------
+# Project: Do Non-Cognitive Skills Affect Employment Outcomes of Youth During the School-to-Work Transition?
+# Organization: SFedU Future Skills Research Lab
+# Objective: Prepare a dataset for the analysis
+# Author:  Garen Avanesian, Valery Egorova
+# Date: 24 June 2024
+#-------------------------------------------------------------------
+
+wave25 <- readRDS(file.path(outData, "rlms_youth25.rds"))
+wave28 <- readRDS(file.path(outData, "rlms_youth28.rds"))
+
+master <- 
+  wave25 %>%
+  full_join(wave28) %>%
+  group_by(idind) %>%
+  mutate(n = 1) %>%
+  mutate(n_total = sum(n))
+
+dim(wave25)
+dim(wave28)
+dim(master)
+
+length(unique(master$idind))
+length(unique(master$idind[master$n_total==2]))
+
+both <- 
+  master %>%
+  filter(n_total == 2) %>%
+  arrange(idind) %>%
+  group_by(idind) 
+  
+  

### 2024-09-28 02_codes/paper_draft.qmd

@@
 The success of school-to-work transition, i.e., if a young person finds a decent job, is determined by various factors such as education, training, as well as supply and demand for qualified workforce on the labor market.
-Those young people who, in addition to being unemployed, also do not study, refer to the most vulnerable group in the school-to-work transition and are defined as not in education, employment, or training (NEET).
-In the recent past, two events amplified vulnerability of youth in terms of labor market outcomes: the Great recession of 2009 and COVID-19 outbreak.
+Those young people who, in addition to being unemployed, also do not study, constitute the most vulnerable group in the school-to-work transition and are defined as not in education, employment, or training (NEET).
+In the recent past, two events amplified vulnerability of youth in terms of labor market outcomes: the Great recession of 2009 and the COVID-19 outbreak.
 The evidence points out that the financial crisis substantially challenged the integration of young people on the labor market all over the world [@tanveerchoudhry2012; @kelly2015; @verick2011; @verd2019; @montalvao2017].
-Young people generally and youth employment particularly were put at the center of the global sustainable development agenda, summarized in Sustainable Development Goal 8 on decent jobs and economic growth.[^1] The global pandemic happened in 2019 amplified the labor market challenges faced by the global community.
+Sone time later, young people generally and youth employment specifically were put at the center of the global sustainable development agenda, summarized in Sustainable Development Goal 8 on decent jobs and economic growth (e.g., SDG 8.6: Promote Youth Employment, Education and Training).
+However, the global pandemic happened in 2019 amplified the labor market challenges faced by the global community.
 It "has heightened the level of uncertainty experienced by many young people worldwide and has complicated the developmental tasks associated with the transition to adulthood, including leaving home, completing education, and obtaining full-time employment" [@allmang2022].
-The initial pandemic estimates suggested that young people were hit more than others by job loss, fell in working hours and substantial drops of income [@ilo2020youth].
-It was estimated that as the result of the pandemic shock in 2020 globally youth employment rates fell by 8.7% in comparison to a 3.7 % drop for adults [@ilo2021].
+The initial pandemic estimates suggested that young people were hit more than others by job loss, fell in working hours and faced substantial drops of income [@ilo2020youth].
+It was estimated that as the result of the pandemic shock in 2020, global youth employment rates fell by 8.7% in comparison to a 3.7 % drop for adults [@ilo2021].
 
-[^1]: Namely, SDG8.6, which asks countries to report the share of youth Not in Education, Employment, or Training.
-
-Though initially the problems of employment and labor market were in the primary interest of economists, the analysis of available academic literature illustrates that in the recent body of work, school-to-work transition was primarily in the focus of psyhologists, sociologists, and then economists, which marks the interdisciplinary nature of the topic [@blokker2023].
-With different disciplines bringing various aspects of employment that were not conventionally regarded by economists, the recent studies point out that non-cognitive skills refer to one of the critical components in school-to-work transition [@lerman2013; @glewwe2017; @zudina2022; @ripamonti2023; @avanesian2024].
-These skills encompass characteristics of personality not directly tied to cognitive abilities, including motivation, self-control, and social skills.
-Research increasingly suggests that non-cognitive skills play a crucial role in determining success in the labor market, influencing both employment opportunities and earning potential [@ferguson2011; @almlund2011a; @almlund2011].
+Though initially the problems of employment and labor market were in the primary interest of economists, the analysis of available academic literature highlights the interdisciplinary nature of the topic, with psychology, sociology, and economics being top 3 academic disciplines to focus on the subject [@blokker2023].
+With different disciplines bringing various determinants of employment that were not conventionally regarded by economists, the recent studies point out that non-cognitive skills refer to one of the critical components in the school-to-work transition [@lerman2013; @glewwe2017; @zudina2022; @ripamonti2023; @avanesian2024].
+These skills encompass characteristics of personality not directly tied to cognitive abilities, including motivation, self-control, social skills, and emotional regulation.
+Research increasingly suggests that non-cognitive skills play a crucial role in determining success in the labor market, influencing both employment opportunities and earning potential in both developed [@ferguson2011; @almlund2011a; @almlund2011] and developing economies [@rozhkova2019; @gimpelson2020].
 
@@
 Further, out of all higher education graduates between 2020-2022 who find a job, for 24% the obtained employment does not match the received profession.
-Finally, the available national data evidence that the labor market in the country demonstrates unprecedented spatial segregation: with the 4% as a country's average unemployment in 2022, it varies in the range between 1.6% in Yamalo-Nenets Autonomous Okrug and 30% in Republic of Ingushetia.[^2]
+Finally, the available national data evidence that the labor market in the country demonstrates unprecedented spatial segregation: with the 4% as a country's average unemployment in 2022, it varies in the range between 1.6% in Yamalo-Nenets Autonomous Okrug and 30% in Republic of Ingushetia.[^1]
 
-[^2]: The Rosstat data were accessed on 24 September 2024, via the following link: <https://rosstat.gov.ru/storage/mediabank/Trud_3_15-s.xlsx>
+[^1]: The Rosstat data were accessed on 24 September 2024, via the following link: <https://rosstat.gov.ru/storage/mediabank/Trud_3_15-s.xlsx>
 
@@
 In order to estimate the effect of non-cognitive skills on employment outcomes, the study relies on the unbalanced panel, where substantial share of observations has only one record.
-In order to address the research questions that drive this investigation, mixed-effects regression model is adopted.
-It's primary advantage refers to the possibility to control endogeneity that occurs due to the unobserved heterogeneity at multiple levels.
+In order to address the research questions that drive this investigation, mixed-effects (multilevel) regression model is adopted [@gelman2006; @mixed-ef2000; @wu2009].
+These models are particularly useful in estimating effects on non-independent, clustered or hierarchical observations and repeated measures [@yang2014].
+Moreover, it's primary advantage refers to the possibility to control endogeneity that occurs due to the unobserved heterogeneity at multiple levels.
 
 First, as the data is longitudinal and XX% of the sample has two records in time, the model adopts random intercept ID to separate between- and within-individual variation in employment and non-cognitive skills.
-Further, the age range of the study is 15-29 in accordance with the ILO definition of the school-to-work transition, but due to differences in baseline probability due to age, it would be the source of endogeneity at the individual level.
-Younger people, due to school enrollment, would be associated with low incidence rates of employment.
+However, to produce unbiased estimates of non-cognitive skills on employment outcomes, the model needs to acknowledge the unobserved heterogeneity in employment due to time-varying characteristics, such as changes in economy, demographics, policy, etc.
+Random intercept for the survey year would isolate the effect of these factors.
+
+Further, the age range of the study refers to 15-29 year-olds in accordance with the ILO definition of the school-to-work transition, but with respect to differences in baseline probability due to age, it would be the source of endogeneity at the individual level.
+Lower incidence rates of employment would be observed amongst younger people enrolled at secondary school or higher education institutions.
 With that regard, the model adopted to estimate the effect of non-cognitive skills on employment will have to isolate the effect of age.
@@
 For that, a second model adopts the random slope terms of non-cognitive skills by the income percentile groups.
+This is also done to acknowledge a confounding role of SES in the relationship between non-cognitive skills and employment, as the random slope term account for the varying relationships between non-cognitive skills and employment outcomes across different SES levels.
+Further, the model automatically estimates a random intercept term for SES by non-cognitive skills as well, which captures the unobserved heterogeneity in the relationship between SES and non-cognitive skills.
 
@@
 
-1\) Quality of data due to the misingness and attrition makes the data potentially highly prone to sample selection bias.
+The study has a number of limitations, both due to the data, as well as estimation strategy.
+First, data missingness and non-response regarding the income-related questions and non-cognitive skills module create substantial challenges for the unbiased estimation, as the smaller set of observations inevitably results in the sample selection bias.
+For participants who took part in both waves of the survey, imputation was used to fill the information available in at least one wave.
 
-Based on the provided methodology section, here are potential research limitations and suggestions to address them.
-I've considered the mentioned data missingness and sample selection bias, as well as other potential issues:
+Attrition bias is also important as the study is based on the unbalanced panel.
+Though the the effect of time-variant characteristics beyond the scope of the analysis, non-random attrition of data also poses implications on the estimation.
+Based on longitudinal design, the analysis does not explicitly address time-varying predictors or dynamic effects, except for controlling for the wave/year of the survey through the random intercept term.
+The further research in this domain could incorporate into the model a number of other time-varying factors that affect both employment outcomes and non-cognitive characteristics.
 
-1.  **Missing Data and Sample Selection Bias:**
-
-    -   *Limitation*: The data suffers from missingness, which can lead to sample selection bias.
-
-    -   *Suggestion*:
-
-        -   Implement multiple imputation techniques to handle missing data.
-This involves creating multiple datasets with imputed values for missing observations and then combining the results from analyzing these datasets.
-R packages like 'mice' or 'Amelia' can be used for this purpose.
-
-        -   Perform sensitivity analyses to assess the impact of different missing data handling strategies on the results.
-
-        -   Consider using inverse probability of censoring weighting (IPCW) or other techniques to address sample selection bias due to missing data.
-
-2.  **Unbalanced Panel and Attrition:**
-
-    -   *Limitation*: The panel is unbalanced, with a substantial share of observations having only one record, and there might be attrition bias.
-
-    -   *Suggestion*:
-
-        -   Investigate the reasons for attrition and compare the characteristics of individuals who dropped out with those who remained in the panel.
-
-        -   Consider using techniques like inverse probability of attrition weighting (IPAW) or shared frailty models to account for attrition bias.
-
-3.  **Endogeneity:**
-
-    -   *Limitation*: The study acknowledges endogeneity due to unobserved heterogeneity at multiple levels (age, region, etc.).
-
-    -   *Suggestion*:
-
-        -   While mixed-effects models help control for endogeneity, consider using instrumental variable (IV) approaches or difference-in-differences (DID) methods for specific research questions where suitable instruments or comparison groups can be identified.
-
-        -   Collect more detailed information in future studies to account for additional unobserved factors.
-
-4.  **Measurement Error:**
-
-    -   *Limitation*: Self-reported data on non-cognitive skills and employment outcomes might be subject to measurement error and social desirability bias.
-
-    -   *Suggestion*:
-
-        -   Consider using objective measures of non-cognitive skills (e.g., cognitive tests, observational studies) or validated scales in future research.
-
-        -   Collect data from multiple sources (e.g., respondents and their parents or teachers) to triangulate information.
-
-5.  **Longitudinal Analysis:**
-
-    -   *Limitation*: The study uses a longitudinal dataset but doesn't explicitly address time-varying covariates or dynamic effects.
-
-    -   *Suggestion*:
-
-        -   Incorporate time-varying covariates in the mixed-effects models to capture changes in individual characteristics over time.
-
-        -   Consider using growth curve models or other advanced longitudinal techniques to analyze dynamic effects.
-
-6.  **Interactions and Non-linear Effects:**
-
-    -   *Limitation*: The study focuses on linear effects and interactions, but non-cognitive skills might have non-linear effects on employment outcomes.
-
-    -   *Suggestion*:
-
-        -   Include quadratic or polynomial terms for non-cognitive skills in the models to capture non-linear effects.
-
-        -   Consider using machine learning techniques or non-parametric approaches to model complex relationships.
-
-By addressing these limitations, you can strengthen the validity and robustness of your research findings.
+Finally, though mixed-effects model estimates the effect of non-cognitive skills on employment accounting for the variation at the different levels (e.g., due to age, region, or SES), it is reached via linear terms.
+Future research that acknowledges non-linearities in personality could help shedding more light on a complex relationship between non-cognitive skills and employment or even a broader set of socio-economic outcomes of an individual.
 

### 2024-09-28 02_codes/0201_data_merge3.R
[code] data prep and regressions

@@ -1,32 +1,0 @@
-#-------------------------------------------------------------------
-# Project: Do Non-Cognitive Skills Affect Employment Outcomes of Youth During the School-to-Work Transition?
-# Organization: SFedU Future Skills Research Lab
-# Objective: Prepare a dataset for the analysis
-# Author:  Garen Avanesian, Valery Egorova
-# Date: 24 June 2024
-#-------------------------------------------------------------------
-
-wave25 <- readRDS(file.path(outData, "rlms_youth25.rds"))
-wave28 <- readRDS(file.path(outData, "rlms_youth28.rds"))
-
-master <- 
-  wave25 %>%
-  full_join(wave28) %>%
-  group_by(idind) %>%
-  mutate(n = 1) %>%
-  mutate(n_total = sum(n))
-
-dim(wave25)
-dim(wave28)
-dim(master)
-
-length(unique(master$idind))
-length(unique(master$idind[master$n_total==2]))
-
-both <- 
-  master %>%
-  filter(n_total == 2) %>%
-  arrange(idind) %>%
-  group_by(idind) 
-  
-  

### 2024-09-28 02_codes/0201_data_prep_2016.R

@@ -29,6 +29,8 @@ data_hh28 <-
                              T ~ as.numeric(NA))) %>%
   select(windex5, family, Income, windex, number_of_fmem, xredid_h, xa3, xid_h, windex, x, region)
 
+# View(data_hh28)
+
 data_hi28 <- 
   hi28 %>%
   select(idind, xredid_i, status, xh5, xj1, x_age, xid_h,  xj161.3y, xj161.3m,
@@ -71,8 +73,10 @@ data_hi28 <-
 ses_data28 <-  
   data_hh28 %>%
   select(xid_h, Income, number_of_fmem , windex, windex5,  x, region) %>%
-  mutate(Income1 = ifelse(is.na(Income), median(Income, na.rm = T), Income),
-         hh_pers = ifelse(is.na(number_of_fmem), median(number_of_fmem, na.rm = T), number_of_fmem)) %>%
+  mutate(Income1 = case_when(is.na(Income) ~ median(Income, na.rm = T), 
+                             TRUE ~ Income),
+         hh_pers = case_when(is.na(number_of_fmem) ~ mean(number_of_fmem, na.rm = T), 
+                             TRUE ~ number_of_fmem)) %>%
   mutate(Income_pc = Income1/number_of_fmem) %>%
   mutate(ses5 = ntile(Income_pc, 5),
          ses10 = ntile(Income_pc, 10),
@@ -125,6 +129,8 @@ data_nc28 <-
          ES = scale(rowMeans(data_nc28[em_st], na.rm = T),center = TRUE, scale = TRUE),
          G = scale(rowMeans(data_nc28[grit], na.rm = T),center = TRUE, scale = TRUE)) %>%
   select(idind,  O, C, E, A, ES, G)
+
+# View(data_nc28)
 
 data_master28 <- 
   data_hi28 %>%

### 2024-09-28 02_codes/0202_data_prep2019.R

@@ -71,15 +71,19 @@ data_hi25 <-
 ses_data25 <-  
   data_hh25 %>%
   select(uid_h, Income, number_of_fmem , windex, windex5,  x, region) %>%
-  mutate(Income1 = ifelse(is.na(Income), median(Income, na.rm = T), Income),
-         hh_pers = ifelse(is.na(number_of_fmem), median(number_of_fmem, na.rm = T), number_of_fmem)) %>%
+  mutate(Income1 = case_when(is.na(Income) ~ median(Income, na.rm = T), 
+                             TRUE ~ Income),
+         hh_pers = case_when(is.na(number_of_fmem) ~ mean(number_of_fmem, na.rm = T), 
+                             TRUE ~ number_of_fmem)) %>%
   mutate(Income_pc = Income1/number_of_fmem) %>%
   mutate(ses5 = ntile(Income_pc, 5),
          ses10 = ntile(Income_pc, 10),
          ses = case_when(ses10 %in% (1:4) ~ "1. Bottom 40%",
                          ses10 %in% (5:9) ~ "2. Middle 50%",
                          ses10 == 10 ~ "3. Top 10%")) %>%
-  select(uid_h, ses5, region, ses10, ses)
+  select(uid_h, ses5, region, ses10, ses, Income_pc, Income, Income1)
+
+View(ses_data25)
 
 openness <- c("o1", "o2", "o3") 
 con <- c("c1", "c2", "c3")
@@ -135,5 +139,8 @@ data_master25 <-
          region = as_factor(region)) %>% 
   select(O, C, E, A, ES, G, ses5, area, sex, age, h_edu, working, seeking_empl, exp, life_satis, 
          self_p_SES, in_education, family_id, salary, region, ses, wave, idind)
+
+# dim(data_hi25)
+# dim(data_master25)
 
 saveRDS(data_master25, file.path(outData, "rlms_youth25.rds"))

### 2024-09-28 02_codes/0202_regressions.R

@@ -1,0 +1,60 @@
+#-------------------------------------------------------------------
+# Project: Do Non-Cognitive Skills Affect Employment Outcomes of Youth During the School-to-Work Transition?
+# Organization: SFedU Future Skills Research Lab
+# Objective: Run the regressions
+# Author:  Garen Avanesian
+# Date: 28 Sept 2024
+#-------------------------------------------------------------------
+
+# m1 <- working ~ age, sex, in_education, ses5, h_edu, area, region, O, C, E, A, ES, 
+# m1 - random intercept by region 
+# m2 - random intercept and random slopes of NCS by ses5
+master <- readRDS(file.path(outData, "model_data.rds"))
+
+m1 <- lmer(working ~ sex + in_education + ses + h_edu + area +  
+               O + C + E + A + ES + 
+               (1|region) + (1|age) + (1|wave) + (1|idind),
+             REML = F, data = master)
+
+summary(m1)
+
+m2 <- lmer(working ~ sex + in_education  + h_edu + area +  
+               O + C + E + A + ES + 
+               (1|region) + (1|age) + (1|wave) + (1|idind) +
+               (1 + O + C + E + A + ES | ses),
+             REML = F, data = master)
+
+mlm_ha_by_ses =
+  coef(m2)$`ses` %>%
+  as.data.frame() %>%
+  rownames_to_column(var = "ses") %>%
+  select(ses, O, C, E, A, ES) %>%
+  gather(Skill, Estimate, -ses) %>%
+  mutate(Estimate = as.numeric(Estimate)*100)
+
+plot_mlm_ha_by_ses =
+  ggplot(mlm_ha_by_ses, aes(Estimate, ses))+
+  geom_point(size = 7, color = "lightblue")+
+  geom_text(aes(label = round(Estimate, 2)), size = 2, color = "black")+
+  #geom_segment( aes(x=0, xend=Estimate, y=SES_Quintile, yend=SES_Quintile), color="skyblue")+
+  geom_vline(xintercept=0, linetype="solid",  color = "black")+
+  theme_bw()+
+  scale_y_discrete(limits=rev)+
+  facet_wrap(Skill~.,)+
+  ylab("")+
+  xlab("")
+
+
+m3 <- lmer(working ~ sex + in_education + ses + area +  
+             O + C + E + A + ES + 
+             (1|region) + (1|age) + (1|wave) + (1|idind) +
[…]

### 2024-09-28 02_codes/0203_data_merge.R

@@ -1,0 +1,38 @@
+#-------------------------------------------------------------------
+# Project: Do Non-Cognitive Skills Affect Employment Outcomes of Youth During the School-to-Work Transition?
+# Organization: SFedU Future Skills Research Lab
+# Objective: Prepare a dataset for the analysis
+# Author:  Garen Avanesian, Valery Egorova
+# Date: 24 June 2024
+#-------------------------------------------------------------------
+
+wave25 <- readRDS(file.path(outData, "rlms_youth25.rds"))
+wave28 <- readRDS(file.path(outData, "rlms_youth28.rds"))
+
+master <- 
+  wave25 %>%
+  full_join(wave28) %>%
+  group_by(idind) %>%
+  mutate(n = 1) %>%
+  mutate(n_total = sum(n)) %>%
+  # SELECT ONLY VARIABLES NEEDED FOR ESTIMATION
+  select(idind, family_id, wave, region, # id vars
+         sex, age, area, ses, ses5, # socio-demographic and hh characteristics
+         working,  in_education, h_edu, # exp,
+         O, C, E, A, ES, # big five traits
+         n, n_total) %>%
+  mutate(in_education = case_when(in_education == 1 ~ in_education,
+                                  TRUE ~ 0)) %>%
+  # get rid of those above 29 who were not in the prev panel
+  mutate(check = case_when(wave == "28" & age >29 & n_total == 1 ~ "drop",
+                           TRUE ~ "keep")) %>%
+  filter(check == "keep")
+
+# dim(master)
+
+# View(master)
+
+saveRDS(master, file.path(outData, "model_data.rds"))
+
+
+

### 2024-09-28 02_codes/0204_descriptive.R

@@ -1,0 +1,12 @@
+#-------------------------------------------------------------------
+# Project: Do Non-Cognitive Skills Affect Employment Outcomes of Youth During the School-to-Work Transition?
+# Organization: SFedU Future Skills Research Lab
+# Objective: Run the descriptive stats
+# Author:  Garen Avanesian
+# Date: 28 Sept 2024
+#-------------------------------------------------------------------
+
+
+# 1. NSC by empl stat with t_test or CI (графики плотности вероятности по 6 хар) T TEST текстом
+# 2. desc table age, sex, area of residence, ses quintile, edu level, in edu - table GA
+# 3. only NCS table by group working (min, max, median, mean, skewness) - obj describe convert to df

### 2024-09-28 02_codes/03_descriptive_analysis.R

@@ -1,3 +1,0 @@
-# 1. NSC by empl stat with t_test or CI (графики плотности вероятности по 6 хар) T TEST текстом
-# 2. desc table age, sex, area of residence, ses quintile, edu level, in edu - table GA
-# 3. only NCS table by group working (min, max, median, mean, skewness) - obj describe convert to df

### 2024-09-28 02_codes/04_regression_analysis.R

@@ -1,54 +1,0 @@
-# m1 <- working ~ age, sex, in_education, ses5, h_edu, area, region, O, C, E, A, ES, 
-# m1 - random intercept by region 
-# m2 - random intercept and random slopes of NCS by ses5
-
-
-mlm1 <- lmer(working ~ sex + in_education + ses5 + h_edu + area +  
-                       O + C + E + A + ES + 
-               + (1|region) + (1|age),
-             REML = F, #control = lmerControl(optimizer = "Nelder_Mead"), 
-             data = data_master)
-
-mlm2 <- lmer(working ~ sex + in_education + ses5 + h_edu + area +  
-               O + C + E + A + ES + self_p_SES  +
-               (1|region) + (1|age) + 
-               (1 + O + C + E + A + ES  | ses5),
-             REML = F, #control = lmerControl(optimizer = "Nelder_Mead"), 
-             data = data_master)
-
-mlm_ha_by_ses =
-  coef(mlm2)$`ses` %>%
-  as.data.frame() %>%
-  rownames_to_column(var = "ses") %>%
-  select(ses, O, C, E, A, ES) %>%
-  gather(Skill, Estimate, -ses) %>%
-  mutate(Estimate = as.numeric(Estimate)*100)
-
-plot_mlm_ha_by_ses =
-  ggplot(mlm_ha_by_ses, aes(Estimate, ses))+
-  geom_point(size = 7, color = "lightblue")+
-  geom_text(aes(label = round(Estimate, 2)), size = 2, color = "black")+
-  #geom_segment( aes(x=0, xend=Estimate, y=SES_Quintile, yend=SES_Quintile), color="skyblue")+
-  geom_vline(xintercept=0, linetype="solid",  color = "black")+
-  theme_bw()+
-  scale_y_discrete(limits=rev)+
-  facet_wrap(Skill~.,)+
-  ylab("")+
-  xlab("")
-
-mlm3 <- lmer(working ~ sex + in_education + ses5 + h_edu + area +  
-               O + C + E + A + ES + self_p_SES + I(O^2) +
-               (1|region) + (1|age) + 
-               (1 + O + C + E + A + ES + I(O^2) | ses5),
-             REML = F, #control = lmerControl(optimizer = "Nelder_Mead"), 
-             data = data_master)
-
-mlm3_scatter =
-  ggpredict(mlm3, terms = c("O","ses5"),
-            type="re") %>% 
[…]

### 2024-10-01 04_manuscript/paper_draft.qmd
[code] data prep and regressions
[code] new merge

--- 2024-09-28 02_codes/paper_draft.qmd
@@
 ---
-title: "Do Non-Cognitive Skills Affect Employment Outcomes of Youth During the School-to-Work Transition?" 
+title: "Do Non-Cognitive Skills Affect Employment of Youth During the School-to-Work Transition?" 
 author: "Garen Avanesian, Southern Federal University"
-format: pdf
+format: 
+  pdf:
+    geometry: 
+      - top=30mm
+      - left=30mm
+      - right=30mm
+    documentclass: article
+    number-sections: true
+    colorlinks: true
+    colorurl: true
 editor: visual
@@
 
-The study analyses the impact of non-cognitive skills on employment outcomes of youth aged 15-29 during the school-to-work transition in Russia.
+The study analyses the impact of non-cognitive skills on employment of youth aged 15-29 during the school-to-work transition in Russia.
 Based on the data of the Russian Longitudinal Monitoring Survey collected in 20XX.
@@
 **JEL Codes:** I24, J24; C51
+
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
+```
+
+```{r wd set up, include=FALSE}
+# set up the path
+path = "/Users/karavan88/Documents/GitHub/youth_empl_ncs"
+
+```
+
+```{r source files, include=FALSE}
+source(file.path(path, "user_profile.R"))
+source(file.path(rCodes, "0200_load_packages.R"))
+# source(file.path(rCodes, "02_descriptive.R"))
+# source(file.path(rCodes, "03_model.R"))
+
+```
 
@@
 While there are many approaches to the definitions of school-to-work transition, their extensive survey is provided in [@blokker2023].
-In this work, I rely on the one that is provided by ILO and defines it "as the passage of a young person (aged 15 to 29 years13) from the end of schooling to the first fixed-term or satisfactory employment" [@ilo2010, p.
+In this work, I rely on the one that is provided by ILO and defines it "as the passage of a young person (aged 15 to 29 years) from the end of schooling to the first fixed-term or satisfactory employment" [@ilo2010, p.
 4].
@@
 Current study explores the effect of non-cognitive skills on employment outcomes of youth aged 15-29 during school-to-work transition in Russia.
-The Russian data highlights that youth has the highest unemployment rates: in 2022, they accounted for 26% for 15-19 year olds, 13.6% for 20-24 year olds, and 5.2% for 25-29 year olds.
-This is compared to 2.8% of unemployment for individuals in the age of 40 to 44 year olds, the lowest value in the sample [@rosstat2024, p.
+The Russian official statistics highlights that youth has the highest unemployment rates: in 2022, they accounted for 26% for 15-19 year olds, 13.6% for 20-24 year olds, and 5.2% for 25-29 year olds.
+This is compared to 2.8% of unemployment for individuals in the age of 40 to 44 year olds, the lowest value by the age cohort breakdown [@rosstat2024, p.
 122].
-Further, out of all higher education graduates between 2020-2022 who find a job, for 24% the obtained employment does not match the received profession.
+Further, out of all higher education graduates between 2020-2022 who found a job, for 24% the obtained employment does not match the received profession.
 Finally, the available national data evidence that the labor market in the country demonstrates unprecedented spatial segregation: with the 4% as a country's average unemployment in 2022, it varies in the range between 1.6% in Yamalo-Nenets Autonomous Okrug and 30% in Republic of Ingushetia.[^1]
@@
 The summary statistics are presented on Table 1.
+
+```{r}
+
+```
 
@@
 
+```{r}
+
+```
+
 Box plot with three facets: employment status, Income percentile, Higher Education
 
-### Effect of Non-Cognitive Skills on Employment Outcomes
+### Effect of Non-Cognitive Skills on Employment
 
@@
 
+```{r}
+
+```
+
 ### Skills Produce Different Effects by Socio-Economic Status
 
+```{r}
+
+```
+
 ### Does Level of Education Increase the Effect of Non-Cognitive Skills?
+
+```{r}
+
+```
 
@@
 
-## Conclusion
+## Conclusions
 

### 2024-10-01 02_codes/0200_load_packages.R

@@ -29,6 +29,7 @@ required_packages <- c("tidyverse",
                        "naniar",
                        "ggpubr",
                        "DescTools", 
+                       "modelsummary",
                        "lme4",
                        "ggeffects", 
                        "sjPlot")

### 2024-10-01 02_codes/0201_data_prep2016.R

@@ -14,6 +14,7 @@ data_hh25 <-
   hh25 %>%
   select(ub1.o, uf14, ua3, UREDID_H, UID_H, region) %>%
   replace_with_na(replace = list(uf14 = c(99999997, 99999998, 99999999))) %>%
+  mutate(hhid25 = UID_H) %>%
   mutate(x = uf14/ub1.o,
          number_of_fmem = ub1.o,
          windex = percent_rank(x),
@@ -27,12 +28,13 @@ data_hh25 <-
                              T ~ as.numeric(NA)),
          uredid_h = UREDID_H,
          uid_h = UID_H) %>%
-  select(windex5, family, Income, windex, number_of_fmem, uredid_h, ua3, uid_h, windex, x, region)
+  select(hhid25, windex5, family, Income, windex, number_of_fmem, 
+         uredid_h, ua3, uid_h, windex, x, region)
 
 data_hi25 <- 
   hi25 %>%
   select(idind, uredid_i, status, uh5, uj1, u_age, uid_h,  uj161.3y, uj161.3m,
-         uj70.2, uj72.1a, uj72.2a, uj72.3a, uj72.4a, uj72.5a, uj72.6a, uj81, 
+         uj70.2, uj72.1a, uj72.2a, uj72.3a, uj72.4a, uj72.5a, uj72.6a, uj81, uj1.1.1,
          u_inwgt, u_diplom, uh3, uj65, uj62, uj322, uj72.171, um3, uj13.2) %>%
   replace_with_na_all(condition ~.x %in% c(99999996, 99999997, 99999998, 99999999)) %>%
   mutate(area = case_when(status %in% c(1,2) ~ "urban",
@@ -64,13 +66,14 @@ data_hi25 <-
          family_id = uh3, 
          seeking_empl = ifelse(uj81 == 1, 1, 0),
          salary = uj13.2) %>%
+  mutate(job_satisfaction = case_when(uj81 %in% c(1,2) ~ 1, TRUE ~ 0 )) %>%
   select(idind, uredid_i, uid_h,  area, gender, age, h_edu, working, seeking_empl, exp, life_satis, 
-         self_p_SES, in_education, family_id, salary, u_inwgt, uj62) %>%
-  filter(age >= 15 & age <= 29)
+          self_p_SES, in_education, family_id, salary, u_inwgt, uj62) #%>%
+  # filter(age >= 15 & age <= 29)
 
 ses_data25 <-  
   data_hh25 %>%
[…]

### 2024-10-01 02_codes/0202_data_prep_2019.R

@@ -14,8 +14,10 @@ hh28        <- read_sav("01_input_data/r28hall_61.sav")
 
 data_hh28 <-
   hh28 %>%
-  select(xb1.o, xf14, xa3, xredid_h, xid_h, region) %>%
+  select(xb1.o, xf14, xa3, xredid_h, xid_h, uid_h, region) %>%
   replace_with_na(replace = list(xf14 = c(99999997, 99999998, 99999999))) %>%
+  mutate(hhid28 = xredid_h,
+         hhid25 = uid_h) %>%
   mutate(x = xf14/xb1.o,
          number_of_fmem = xb1.o,
          windex = percent_rank(x),
@@ -27,13 +29,14 @@ data_hh28 <-
                              windex >= 0.61 & windex <= 0.80 ~ 4,
                              windex > 0.8  ~ 5,
                              T ~ as.numeric(NA))) %>%
-  select(windex5, family, Income, windex, number_of_fmem, xredid_h, xa3, xid_h, windex, x, region)
+  select(hhid28, hhid25, windex5, family, Income, windex, number_of_fmem, 
+         xredid_h, xa3, xid_h, windex, x, region)
 
 # View(data_hh28)
 
 data_hi28 <- 
   hi28 %>%
-  select(idind, xredid_i, status, xh5, xj1, x_age, xid_h,  xj161.3y, xj161.3m,
+  select(idind, xredid_i, status, xh5, xj1, x_age, xid_h,  xj161.3y, xj161.3m, xj1.1.1,
          xj70.2, xj72.1a, xj72.2a, xj72.3a, xj72.4a, xj72.5a, xj72.6a, xj81, 
          x_inwgt, x_diplom, xh3, xj65, xj62, xj322, xj72.171, xm3, xj13.2) %>%
   replace_with_na_all(condition ~.x %in% c(99999996, 99999997, 99999998, 99999999)) %>%
@@ -66,13 +69,14 @@ data_hi28 <-
          family_id = xh3, 
          seeking_empl = ifelse(xj81 == 1, 1, 0),
          salary = xj13.2) %>%
+  mutate(job_satisfaction = case_when(xj1.1.1 %in% c(1,2) ~ 1, TRUE ~ 0 )) %>%
   select(idind, xredid_i, xid_h,  area, gender, age, h_edu, working, seeking_empl, exp, life_satis, 
-         self_p_SES, in_education, family_id, salary, x_inwgt, xj62) %>%
-  filter(age >= 15 & age <= 34)
[…]

### 2024-10-01 02_codes/0203_data_merge.R

@@ -12,14 +12,18 @@ wave28 <- readRDS(file.path(outData, "rlms_youth28.rds"))
 master <- 
   wave25 %>%
   full_join(wave28) %>%
+  filter(age >= 15 & age <= 34) %>%
+  mutate(hh_id_fin = case_when(!is.na(hhid28) ~ hhid28,
+                           TRUE ~ as.numeric(paste0("25", hhid25)))) %>%
+  #group_by(hh_id_fin) %>%
   group_by(idind) %>%
   mutate(n = 1) %>%
   mutate(n_total = sum(n)) %>%
   # SELECT ONLY VARIABLES NEEDED FOR ESTIMATION
-  select(idind, family_id, wave, region, # id vars
+  select(idind, family_id, wave, region, hhid25, hhid28, hh_id_fin, # id vars
          sex, age, area, ses, ses5, # socio-demographic and hh characteristics
          working,  in_education, h_edu, # exp,
-         O, C, E, A, ES, # big five traits
+         O, C, E, A, ES, G, # big five traits
          n, n_total) %>%
   mutate(in_education = case_when(in_education == 1 ~ in_education,
                                   TRUE ~ 0)) %>%
@@ -33,6 +37,3 @@ master <-
 # View(master)
 
 saveRDS(master, file.path(outData, "model_data.rds"))
-
-
-

### 2024-10-01 02_codes/0205_regressions.R

@@ -12,23 +12,25 @@
 master <- readRDS(file.path(outData, "model_data.rds"))
 
 m1 <- lmer(working ~ sex + in_education + ses + h_edu + area +  
-               O + C + E + A + ES + 
+               O + C + E + A + ES + G +
                (1|region) + (1|age) + (1|wave) + (1|idind),
              REML = F, data = master)
 
 summary(m1)
 
 m2 <- lmer(working ~ sex + in_education  + h_edu + area +  
-               O + C + E + A + ES + 
+               O + C + E + A + ES + G +
                (1|region) + (1|age) + (1|wave) + (1|idind) +
-               (1 + O + C + E + A + ES | ses),
+               (1 + O + C + E + A + ES + G | ses),
              REML = F, data = master)
+
+summary(m2)
 
 mlm_ha_by_ses =
   coef(m2)$`ses` %>%
   as.data.frame() %>%
   rownames_to_column(var = "ses") %>%
-  select(ses, O, C, E, A, ES) %>%
+  select(ses, O, C, E, A, ES, G) %>%
   gather(Skill, Estimate, -ses) %>%
   mutate(Estimate = as.numeric(Estimate)*100)
 
